---
title: "Final Logistic"
author: "Dara Zirlin"
date: "July 23, 2017"
output: html_document
---

# Load needed packages and csv files
```{r}

suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(mosaic))




setwd('C:/Users/DaraZ/Documents/IMA/tennis')
train <- read.csv('train_dataset20002013_1Jul15.csv')

valid<- read.csv('valid_dataset20142015_1Jul15.csv')

test<-read.csv('test_dataset20162017.csv')
all<-rbind(train,valid,test)

train<-train%>%mutate_if(is.factor,as.character)
valid<-valid%>%mutate_if(is.factor,as.character)
test<-test%>%mutate_if(is.factor,as.character)

train<-na.omit(train)
valid<-na.omit(valid)
test<-na.omit(test)

valid[valid==""]<-NA
```


# Needed functions
```{r}

#This function outputs player as player1
p1constant <- function(traindata,player){
  p1data <- traindata %>% filter(Player_1 == player | Player_2 == player)

for(p in  1:nrow(p1data)){  
  if (p1data[p,]$Player_2 == player){
  a=p1data[p,]$Player_1
  b=p1data[p,]$Player_2
  
  p1data[p,]$Player_1=b
  p1data[p,]$Player_2=a
  
  if (p1data[p,]$won == 0){p1data[p,]$won=1} else {p1data[p,]$won=0}
 
  
 colns1 <- c('age_diff', 'rankpts_diff', 'ranks_diff','ratio_ace_diff','ratio_1stIn_diff','ratio_1stWon_diff','ratio_2ndWon_diff')
  
 p1data[p,colns1 ]=-p1data[p,colns1 ]
  }
}
  return(p1data)
}






```

# specific player data
```{r}
roger<-p1constant(train,"Roger Federer")
roger_valid<-p1constant(valid,"Roger Federer")
roger_test<-p1constant(test,"Roger Federer")


andy<-p1constant(train,"Andy Murray")
andy_valid<-p1constant(valid,"Andy Murray")
andy_test<-p1constant(test,"Andy Murray")



nadal<-p1constant(train,"Rafael Nadal")
nadal_valid<-p1constant(valid,"Rafael Nadal")
nadal_test<-p1constant(test,"Rafael Nadal")

```

# Rank comparison on the general model
```{r}

# Store the predicition of rank comparison
predict_joe<-valid%>%mutate(predict=(ranks_diff<=0))

# Calculate the proportion predicted correctly in the validation set
1-sum((predict_joe$predict-valid$won)^2)/length(valid$won)



```



# Best Logistic Model on the General Dataset
```{r}

#Fit the model
mod <- glm(won~ranks_diff+rankpts_diff+ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+avg_brkpt_converted+surface, train,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)

# Store the predicition of the match
prediction<-prediction%>%mutate(probPrediction=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff))

#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])




# turn winning into a factor

prediction$won<-as.factor(prediction$won)


# plot the probability of winning vs if a player won
ggplot(prediction, aes(x=probPrediction, fill=won)) +
  geom_density(alpha=.5)+labs(x="Prdicted Probability",title="General Population Predicition")+scale_fill_manual(labels = c("lost", "won"),values=c("cyan","pink")) 








```



#  Best Logisitic Model on the General dataset not using rank
```{r}

# fit the model
mod <- glm(won~ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+avg_brkpt_converted+surface, train,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)

#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```

# Rank Comparison Model on Roger Federer
```{r}


# Store the predicition of rank comparison
predict_joe<-roger_valid%>%mutate(predict=(ranks_diff<=0))

# Calculate the proportion predicted correctly in the validation set
1-sum((predict_joe$predict-roger_valid$won)^2)/length(roger_valid$won)




```





#  Roger Federer  Logistic Model with rank
```{r}

# fit the model
mod <- glm(won~rankpts_diff+ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+age_diff, roger,family="binomial")

summary(mod)

# create a function based off the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-roger_valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)

#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```



#   Roger Federed Logistic Model without ranks
```{r}

# fit the model
mod <- glm(won~ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+age_diff, roger,family="binomial")

summary(mod)

# create a function based off the model
probFun <- makeFun(mod)


#Store the predicted probabilities of a player winning the match from the validation set
prediction<-roger_valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)



#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```


# Rank Comparison Model of Andy Murray
```{r}



# Store the predicition of rank comparison
predict_joe<-andy_valid%>%mutate(predict=(ranks_diff<=0))

# Calculate the proportion predicted correctly in the validation set
1-sum((predict_joe$predict-andy_valid$won)^2)/length(andy_valid$won)






```

#   Andy Murray Logistic Model  with rank
```{r}

#fit the model
mod <- glm(won~rankpts_diff+ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff, andy,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)


#Store the predicted probabilities of a player winning the match from the validation set
prediction<-andy_valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)


#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```



#   Andy Murray Logistic Model  without rank
```{r}

#fit the model
mod <- glm(won~ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+hands+age_diff, andy,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-andy_valid%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)



#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```

# Rank Comparison Model on Rafael Nadal
```{r}


# Store the predicition of rank comparison
predict_joe<-nadal_valid%>%mutate(predict=(ranks_diff<=0))

# Calculate the proportion predicted correctly in the validation set
1-sum((predict_joe$predict-nadal_valid$won)^2)/length(nadal_valid$won)

```

#  Rafael Nadal Logistic Model with ranks
```{r}

#fit the model
mod <- glm(won~rankpts_diff+surface+age_diff, nadal,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-nadal_test%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)



#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```





#  Rafael Nadal Logistic Model without ranks
```{r}

#fit the model
mod <- glm(won~ratio_1stWon_diff+ratio_2ndWon_diff+ratio_ace_diff+surface+age_diff, nadal,family="binomial")

summary(mod)

# create a function based off of the model
probFun <- makeFun(mod)

#Store the predicted probabilities of a player winning the match from the validation set
prediction<-nadal_test%>%mutate(predict=probFun(ratio_1stWon_diff=ratio_1stWon_diff,avg_brkpt_converted=avg_brkpt_converted,ratio_2ndWon_diff=ratio_2ndWon_diff,ranks_diff=ranks_diff,rankpts_diff=rankpts_diff,log_rankpts_diff=log_rankpts_diff,ratio_1stIn_diff=ratio_1stIn_diff,surface=surface,hands=hands,ratio_ace_diff=ratio_ace_diff,age_diff=age_diff)>.5)



#Calculate the proportion predicted correctly in the validation set
1-sum((prediction$predict[!is.na(prediction$predict)]-prediction$won[!is.na(prediction$predict)])^2)/length(prediction$predict[!is.na(prediction$predict)])

```