---
title: "Classification Trees - Tennis - Day 2"
author: "Missy Gaddy"
date: "7/15/2017"
output: html_document
---

** Models without Rank - all Players 

Get data and package
```{r}
#install.packages("tree")
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tree))
suppressPackageStartupMessages(library(ggplot2))
#setwd("/Users/checkout/Google Drive/Summer 2017/Math to Industry Bootcamp/Project 1 - Tennis/IMA_BootCamp_Project1")

# Loading training and validation data for all players (with breakpoint info)
data_train_all <- read.csv("train_dataset20002013_1Jul15.csv")
data_validate_all <- read.csv("valid_dataset20142015_1Jul15.csv")


```


Adding a row with the Bradley-Terry probability:
```{r}

data_raw <- read.csv('raw_dataset20002013.csv')

data_raw <- data_raw%>% mutate(winner_p1 = winner_rank/(winner_rank + loser_rank), loser_p1 = loser_rank/(winner_rank+loser_rank)) %>% mutate(BTpred = ifelse(winner_p1 >loser_p1, 0,1))

sum(na.omit(data_raw$BTpred))/length(na.omit(data_raw$BTpred))
```


```{r}
# make "won" into a factor for datasets 

data_train_all <- data_train_all %>% mutate(won_factor = factor(won))
data_validate_all <- data_validate_all %>% mutate(won_factor = factor(won))


```

** Model with all factors minus rank, rankpts, and log_rankpts
```{r}
tree.tennis_all_noRank = tree(won_factor ~ surface + hands  + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted, data=data_train_all)
summary(tree.tennis_all_noRank)
tree.tennis_all_noRank

# Make prediction
prediction <- predict(tree.tennis_all_noRank, data_validate_all,type = "class") # gives the probability 
table(prediction,data_validate_all$won_factor)

# Plotting the tree
plot(tree.tennis_all_noRank)
text(tree.tennis_all_noRank, pretty = 0)
```


```{r}
(803+2207)/(803+2207+680+2084)
```

Looking at if we should prune. Here is the number of nodes vs. the cross-validation error:
```{r}
cv.tennis_all_noRank = cv.tree(tree.tennis_all_noRank, FUN = prune.misclass)
plot(cv.tennis_all_noRank$size, cv.tennis_all_noRank$dev, type="b")
```

Now pruning down to 8 nodes

```{r}
prune.tennis_all_noRank = prune.misclass(tree.tennis_all_noRank, best = 8)
prediction_pruned_all_noRank <- predict(prune.tennis_all_noRank, data_validate_all,type = "class")
table(prediction_pruned_all_noRank,data_validate_all$won_factor)

# Plotting the pruned tree
plot(prune.tennis_all_noRank)
text(prune.tennis_all_noRank, pretty = 0)
```
The error is the same. :(








--------------------------------------------------

** Models with ALL predictors

There are 11 predictors with our data

```{r}

tree.tennis_all = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ranks_diff + log_rankpts_diff + rankpts_diff, data=data_train_all)
summary(tree.tennis_all)
tree.tennis_all

plot(tree.tennis_all)
text(tree.tennis_all, pretty = 0)

cv.tennis_all = cv.tree(tree.tennis_all, FUN = prune.misclass)
plot(cv.tennis_all$size, cv.tennis_all$dev, type="b")

```



Now making prediction:
```{r}
prediction_all <- predict(tree.tennis_all, data_validate_all,type = "class") # gives the probability for each class
#head(prediction)
table(prediction_all,data_validate_all$won_factor)
```

So this is exactly a 50-50 split. Now let's prune:
```{r}
prune.tennis_all = prune.misclass(tree.tennis_all, best = 4)
prediction_pruned_all <- predict(prune.tennis_all, data_validate_all,type = "class")
table(prediction_pruned_all,data_validate_all$won_factor)

plot(prune.tennis_all)
text(prune.tennis_all, pretty=0)
```

Now the prediction accuracy is
```{r}
(18+2861)/(18+2861+26+2869)
```





----------------------------------------------------

** Models with Just Rank_diff, log_rankpts, rankpts - all Players

```{r}
tree.tennis_all_onlyRank = tree(won_factor ~ ranks_diff + log_rankpts_diff + rankpts_diff , data=data_train_all)
summary(tree.tennis_all_onlyRank)
tree.tennis_all_onlyRank
plot(tree.tennis_all_onlyRank)
text(tree.tennis_all_onlyRank, pretty = 0)
cv.tennis_all_onlyRank = cv.tree(tree.tennis_all_onlyRank, FUN = prune.misclass)
plot(cv.tennis_all_onlyRank$size, cv.tennis_all_onlyRank$dev, type="b")
```

Prediction:
```{r}
prediction_onlyRank <- predict(tree.tennis_all_onlyRank, data_validate_all,type = "class") # gives the probability for each class
#head(prediction)
table(prediction_onlyRank,data_validate_all$won_factor)
```
This is the prediction accuracy for the model with rank_diff, rank_pts, and log(rank_pts)
```{r}
(58+2812)/(58+2812+75+2829)
```

From the table above, I can prune this to 4 nodes.
```{r}
prune.tennis_all_onlyRank = prune.misclass(tree.tennis_all_onlyRank, best = 4)
prediction_pruned_all_onlyRank <- predict(prune.tennis_all_onlyRank, data_validate_all,type = "class")
table(prediction_pruned_all_onlyRank,data_validate_all$won_factor)

plot(prune.tennis_all_onlyRank)
text(prune.tennis_all_onlyRank, pretty=0)
```

```{r}
(109+2758)/(109+2758+2278+129)
```






------------------------------------------------------

** Model With only rank_diff
```{r}
tree.tennis_all_onlyRankDiff = tree(won_factor ~ ranks_diff, data=data_train_all)
summary(tree.tennis_all_onlyRankDiff)
tree.tennis_all_onlyRankDiff
plot(tree.tennis_all_onlyRankDiff)
text(tree.tennis_all_onlyRankDiff, pretty = 0)
```

Prediction 
```{r}
prediction_onlyRankDiff <- predict(tree.tennis_all_onlyRank, data_validate_all,type = "class") # gives the probability for each class
#head(prediction)
table(prediction_onlyRankDiff,data_validate_all$won_factor)
```


This is the prediction accuracy for the model with just rank_diff
```{r}
(1959+1948)/(1959+1948+939+928)
```

So the models with only rank_diff get 67% accuracy.







---------------------------------------------------------------

** Model with Breakpoints, 1st won, and 2nd won

```{r}
tree.tennis_all_Bkpt1st2nd = tree(won_factor ~ avg_brkpt_converted + ratio_1stWon_diff + ratio_2ndWon_diff, data=data_train_all)
summary(tree.tennis_all_Bkpt1st2nd)
tree.tennis_all_Bkpt1st2nd
plot(tree.tennis_all_Bkpt1st2nd)
text(tree.tennis_all_Bkpt1st2nd, pretty = 0)

cv.tennis_all_Bkpt1st2nd = cv.tree(tree.tennis_all_Bkpt1st2nd, FUN = prune.misclass)
plot(cv.tennis_all_Bkpt1st2nd$size, cv.tennis_all_Bkpt1st2nd$dev, type="b")
```

```{r}
prediction_Bkpt1st2nd <- predict(tree.tennis_all_Bkpt1st2nd, data_validate_all,type = "class") # gives the probability for each class
#head(prediction)
table(prediction_Bkpt1st2nd,data_validate_all$won_factor)
```

```{r}
(803+2207)/(803+2207+2084+680)

```

Now I will prune down to 8 nodes:
```{r}
prune.tennis_all_Bkpt1st2nd = prune.misclass(tree.tennis_all_Bkpt1st2nd, best = 8)
prediction_pruned_all_Bkpt1st2nd <- predict(prune.tennis_all_Bkpt1st2nd, data_validate_all,type = "class")
table(prediction_pruned_all_Bkpt1st2nd,data_validate_all$won_factor)

plot(prune.tennis_all_Bkpt1st2nd)
text(prune.tennis_all_Bkpt1st2nd, pretty=0)
```

Now the prediction with the pruned tree is

```{r}
(803+2207)/(803+2207+680+2084)
```






--------------------------------------------------------
** Model with 1stWon only

```{r}
# Make tree
tree.tennis_all_1stWon = tree(won_factor ~ ratio_1stWon_diff, data=data_train_all)
summary(tree.tennis_all_1stWon)
tree.tennis_all_1stWon

# plot tree
plot(tree.tennis_all_1stWon)
text(tree.tennis_all_1stWon, pretty = 0)


# Make prediction on unpruned tree
prediction_1stWon <- predict(tree.tennis_all_1stWon, data_validate_all,type = "class") # gives the
table(prediction_1stWon,data_validate_all$won_factor)

# Look at potential to prune
cv.tennis_all_1stWon = cv.tree(tree.tennis_all_1stWon, FUN = prune.misclass)
plot(cv.tennis_all_1stWon$size, cv.tennis_all_1stWon$dev, type="b")
```

Prediction accuracy with unpruned tree
```{r}
(1508+1833)/(1508+1833+1379+1054)
```

Silly, you can't prune a tree with 2 nodes...








-------------------------------------------------------------------
** Model with only Breakpoints

```{r}
# Make tree
tree.tennis_all_bkpt = tree(won_factor ~ avg_brkpt_converted, data=data_train_all)
summary(tree.tennis_all_bkpt)
tree.tennis_all_bkpt

# plot tree
plot(tree.tennis_all_bkpt)
text(tree.tennis_all_bkpt, pretty = 0)


# Make prediction on unpruned tree
prediction_bkpt <- predict(tree.tennis_all_bkpt, data_validate_all,type = "class") # gives the
table(prediction_bkpt,data_validate_all$won_factor)


```

Whoops, it doesn't think bkpt is significant enough to branch on.







----------------------------------------------------------------
** Model with Breakpt, 1stWon, 2ndWon, Aces, Age

```{r}
# Make tree
tree.tennis_all_B12AA = tree(won_factor ~ age_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ratio_ace_diff, data=data_train_all)
summary(tree.tennis_all_B12AA)
tree.tennis_all_B12AA

# plot tree
plot(tree.tennis_all_B12AA)
text(tree.tennis_all_B12AA, pretty = 0)


# Make prediction on unpruned tree
prediction_B12AA <- predict(tree.tennis_all_B12AA, data_validate_all,type = "class") # gives the
table(prediction_B12AA,data_validate_all$won_factor)

# Look at potential to prune
cv.tennis_all_B12AA = cv.tree(tree.tennis_all_B12AA, FUN = prune.misclass)
plot(cv.tennis_all_B12AA$size, cv.tennis_all_B12AA$dev, type="b")
```

Prediction accuracy:
```{r}
(803+2207)/(803+2207+2084+680)
```




-----------------------------------------------------------------
** Individual Players

So there may be predictors that have more weight for certain players. Like surface and stuff

*** Roger Federer

Loading Roger's data
```{r}
data_Federer_train <- read.csv("train_dataset20002013_1Jul15_roger.csv")
#data_Federer_validate <- read.csv("valid_dataset20142015_roger.csv") # does not include breakpoint info
data_Federer_validate<-p1constant(data_validate_all,"Roger Federer") # this function is inside GetAnotherPlayerData.R

# make won into a factor
data_Federer_train <- data_Federer_train %>% mutate(won_factor = factor(won))
data_Federer_validate <- data_Federer_validate %>% mutate(won_factor = factor(won))

# tree with all predictors
tree.tennis_Federer = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ranks_diff + log_rankpts_diff, data=data_Federer_train)

# tree without rank predictors, except keeping ranks_diff
#tree.tennis_Federer = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ranks_diff, data=data_Federer_train)

# tree without rank predictors
#tree.tennis_Federer = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted, data=data_Federer_train)

summary(tree.tennis_Federer)
tree.tennis_Federer

plot(tree.tennis_Federer)
text(tree.tennis_Federer, pretty = 0)

# Make prediction on unpruned tree
prediction_Federer <- predict(tree.tennis_Federer, data_Federer_validate,type = "class") # gives the
table(prediction_Federer,data_Federer_validate$won_factor)

cv.tennis_Federer = cv.tree(tree.tennis_Federer, FUN = prune.misclass)
plot(cv.tennis_Federer$size, cv.tennis_Federer$dev, type="b")


```

```{r}
(115+10)/(115+10+13+19)
```


*** Rafael Nadal

Loading Nadal's data
```{r}
#data_train_all <- read.csv("train_dataset20002013_1Jul15.csv")
#data_Federer_validate <- read.csv("valid_dataset20142015_roger.csv") # does not include breakpoint info
data_Nadal_train <- p1constant(data_train_all, "Rafael Nadal")
data_Nadal_validate<-p1constant(data_validate_all,"Rafael Nadal") # this function is inside GetAnotherPlayerData.R

# make won into a factor
data_Nadal_train <- data_Nadal_train %>% mutate(won_factor = factor(won))
data_Nadal_validate <- data_Nadal_validate %>% mutate(won_factor = factor(won))

# tree with all predictors
tree.tennis_Nadal = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ranks_diff + log_rankpts_diff + rankpts_diff, data=data_Nadal_train)

# tree without rank predictors, except keeping ranks_diff
#tree.tennis_Federer = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted + ranks_diff, data=data_Federer_train)

# tree without rank predictors
#tree.tennis_Nadal = tree(won_factor ~ surface + hands + age_diff + ratio_ace_diff + ratio_1stIn_diff + ratio_1stWon_diff + ratio_2ndWon_diff + avg_brkpt_converted, data=data_Nadal_train)

summary(tree.tennis_Nadal)
tree.tennis_Nadal

plot(tree.tennis_Nadal)
text(tree.tennis_Nadal, pretty = 0)

# Make prediction on unpruned tree
#prediction_Federer <- predict(tree.tennis_Federer, data_Federer_validate,type = "class") # gives the
#table(prediction_Federer,data_Federer_validate$won_factor)

#cv.tennis_Federer = cv.tree(tree.tennis_Federer, FUN = prune.misclass)
#plot(cv.tennis_Federer$size, cv.tennis_Federer$dev, type="b")


```


** Predicting tomorrow's match

```{r}
#federer_newest <- test_dataset20162017 %>% filter(Player_1 == "Roger Federer" | Player_2 == "Roger Federer")
data_raw_newest <- read.csv("raw_testdataset20162017.csv")
federer_mostRecent1 <- data_raw_newest %>% filter(tourney_date == 20170116 & winner_name == "Roger Federer")
federer_mostRecent2 <- test_dataset20162017 %>% filter(tourney_date == 20170116 & (Player_1 == "Roger Federer" | Player_2 == "Roger Federer"))

cilic_mostRecent <- test_dataset20162017 %>% filter(Player_1 == "Marin Cilic" | Player_2 == "Marin Cilic")

cilic_playingWawrinka <- cilic_mostRecent[70,]
federer_playingWawrinka <- federer_mostRecent2[6,]

agediff_federer_cilic <- cilic_playingWawrinka$age_diff - federer_playingWawrinka$age_diff # cilic minus federer
ranksdiff_federer_cilic <- cilic_playingWawrinka$ranks_diff - federer_playingWawrinka$ranks_diff
print(ranksdiff_federer_cilic) # cilic minus federer

ratioacediff_federer_cilic <- cilic_playingWawrinka$ratio_ace_diff - federer_playingWawrinka$ratio_ace_diff

names(data_Federer_train)

```


Putting together data row
```{r}
numeric_federer_playingWawrinka <- data.frame(federer_playingWawrinka$age_diff, federer_playingWawrinka$log_rankpts_diff, federer_playingWawrinka$rankpts_diff, federer_playingWawrinka$ranks_diff, federer_playingWawrinka$ratio_ace_diff, federer_playingWawrinka$ratio_1stIn_diff, federer_playingWawrinka$ratio_1stWon_diff, federer_playingWawrinka$ratio_2ndWon_diff, federer_playingWawrinka$avg_brkpt_converted)


numeric_cilic_playingWawrinka <- data.frame(cilic_playingWawrinka$age_diff, cilic_playingWawrinka$log_rankpts_diff, cilic_playingWawrinka$rankpts_diff, cilic_playingWawrinka$ranks_diff, cilic_playingWawrinka$ratio_ace_diff, cilic_playingWawrinka$ratio_1stIn_diff, cilic_playingWawrinka$ratio_1stWon_diff, cilic_playingWawrinka$ratio_2ndWon_diff, cilic_playingWawrinka$avg_brkpt_converted)

difference <- numeric_federer_playingWawrinka - numeric_cilic_playingWawrinka
colnames(difference) <- c("age_diff", "log_rankpts_diff", "rankpts_diff", "ranks_diff", "ratio_ace_diff","ratio_1stIn_diff","ratio_1stWon_diff", "ratio_2ndWon_diff", "avg_brkpt_converted")

gameTomorrow <- data.frame(X = 100, tourney_date = 20170716, match_num= 100, Player_1 = "Roger Federer", Player_2 = "Marin Cilic", surface = "Grass", hands = TRUE, difference, won = 0, won_factor = 0)

 
```

```{r}
# Predicting wimbledon outcome without rank
prediction_wimbledon_noRank <- predict(tree.tennis_all_noRank, gameTomorrow,type = "class") # gives the probability 
prediction_wimbledon_1stWon <- predict(tree.tennis_all_1stWon, gameTomorrow, type="class")
prediction_wimbledon_onlyRankDiff <- predict(tree.tennis_all_onlyRankDiff, gameTomorrow, type = "class")
prediction_wimbledon_all <- predict(tree.tennis_all, gameTomorrow, type="class")

prediction_wimbledon_noRank
prediction_wimbledon_1stWon
prediction_wimbledon_onlyRankDiff
prediction_wimbledon_all
```
